---
title: "final_project"
output: html_document
---

```{r Needed Files from the MIMIC Data}
# - chartevents.csv-003
# from mimic 2 hosp folder
# - patients.csv
# - admissions.csv
# - diagnoses_icd.csv
# from mimic 2 icu folder
# - icustays.csv
# - d_items.csv
```

```{r 1 Data Loading}
library(data.table)

patients <- fread("/Users/brycen.izu/stats c160/mimic-iv-2.2 2/hosp/patients.csv")
admissions <- fread("/Users/brycen.izu/stats c160/mimic-iv-2.2 2/hosp/admissions.csv")
diagnoses <- fread("/Users/brycen.izu/stats c160/mimic-iv-2.2 2/hosp/diagnoses_icd.csv")
icustays <- fread("/Users/brycen.izu/stats c160/mimic-iv-2.2 2/icu/icustays.csv")

# Star with icustays

data <- icustays

# Merge admissions
data <- merge(
  data,
  admissions[, .(subject_id, hadm_id, hospital_expire_flag, admission_type)],
  by = c("subject_id", "hadm_id")
)

data[, death := hospital_expire_flag == 1]

# Merge patients
data <- merge(data, patients[, .(subject_id, anchor_age, gender)],
              by = "subject_id")
data[, age := anchor_age]

# Merge diagnoses
# install.packages("comorbidity")
library(comorbidity)

diag9  <- diagnoses[icd_version == 9]
diag10 <- diagnoses[icd_version == 10]

elix9 <- comorbidity(
  x = diag9,
  id = "hadm_id",
  code = "icd_code",
  map = "elixhauser_icd9_quan",
  assign0 = TRUE
)

elix10 <- comorbidity(
  x = diag10,
  id = "hadm_id",
  code = "icd_code",
  map = "elixhauser_icd10_quan",
  assign0 = TRUE
)

comorb_names <- union(
  setdiff(names(elix9), "hadm_id"),
  setdiff(names(elix10), "hadm_id")
)

setDT(elix9)
setDT(elix10)

elix <- merge(elix9, elix10, by = "hadm_id", all = TRUE, suffixes = c("_9", "_10"))
setDT(elix)

for (nm in comorb_names) {
  c9 <- paste0(nm, "_9")
  c10 <- paste0(nm, "_10")
  
  if (!c9 %in% names(elix)) elix[, (c9) := 0]
  if (!c10 %in% names(elix)) elix[, (c10) := 0]
  
  elix[, (nm) := pmax(get(c9), get(c10), na.rm = TRUE)]
}

elix <- elix[, c("hadm_id", comorb_names), with = FALSE]

data <- merge(data, elix, by = "hadm_id", all.x = TRUE)

for (nm in comorb_names) {
  data[is.na(get(nm)), (nm) := 0]
}

# Merge d_items
items <- fread("/Users/brycen.izu/stats c160/mimic-iv-2.2 2/icu/d_items.csv")
items[grep("height|weight", label, ignore.case=TRUE)]

# Merge chartevents
weight_id <- c(226512)
height_id <- c(226730)

chart <- fread(
  "chartevents.csv-003",
  select = c("subject_id","hadm_id","itemid","valuenum")
)

hw <- chart[itemid %in% c(weight_id, height_id)]

weight <- hw[itemid %in% weight_id,
             .(subject_id, hadm_id, weight = valuenum)]

height <- hw[itemid %in% height_id,
             .(subject_id, hadm_id, height = valuenum)]

weight <- weight[weight >= 30 & weight <= 300]
height <- height[height >= 120 & height <= 230]

weight_one <- weight[!is.na(weight),
  .(weight = median(weight, na.rm = TRUE)),
  by = .(subject_id, hadm_id)
]

height_one <- height[!is.na(height),
  .(height = median(height, na.rm = TRUE)),
  by = .(subject_id, hadm_id)
]

bmi_data <- merge(weight_one, height_one, by = c("subject_id", "hadm_id"))

bmi_data[, height_m := height / 100]
bmi_data[, bmi_all := weight / (height_m^2)]

data <- merge(
  data,
  bmi_data[, .(subject_id, hadm_id, bmi_all)],
  by = c("subject_id", "hadm_id"),
  all.x = TRUE
)

setorder(data, subject_id, hadm_id, intime)
data <- data[, .SD[1], by = .(subject_id, hadm_id)]
```

```{r 2 Selecting Cohort & Generating Data}
data <- data[complete.cases(data[, .(death, bmi_all, age)])]

data <- data[bmi_all >= 18.5 & bmi_all < 35]

# Add Treatment variable
data[, treat := as.integer(bmi_all >= 25)]

comorb_vars <- intersect(comorb_names, names(data))

for (var in comorb_vars) {
  data[is.na(get(var)), (var) := 0]
}

keep_vars <- c("bmi_all", "treat", "death", "age", "gender")
data <- data[complete.cases(data[, ..keep_vars])]

# Prepare sex and death for modeling
data[, sex := fifelse(gender %in% c("M", "Male"), 1L, 0L)]
data[, death := as.integer(death)]

data[, admission_type := as.factor(admission_type)]
adm_mat <- model.matrix(~ admission_type - 1, data = data)

data[, c("obes", "wloss") := NULL]

comorb_vars <- setdiff(comorb_vars, c("obes", "wloss"))

data[, diab := pmax(diabunc, diabc)]
data[, c("diabunc", "diabc") := NULL]

comorb_vars <- setdiff(comorb_vars, c("diabunc", "diabc"))
comorb_vars <- union(comorb_vars, "diab")

data[, comorbidity_burden := rowSums(.SD), .SDcols = comorb_vars]
```

```{r 3 Inspecting Heterogeneity}
library(grf)

covars_num <- c("age", "sex", comorb_vars)

X <- cbind(as.matrix(data[, ..covars_num]), adm_mat)
Y <- as.numeric(data$death)
W <- data$treat

crf <- causal_forest(
  X = X,
  Y = Y,
  W = W
)

data[, cate := predict(crf)$predictions]
```

```{r 3.1 TE Measure}
ate <- mean(data$cate)
ate

round(ate, 4)
```

The treatment decreases mortality probability by 0.0111 on average.

```{r 3.2 ETT}
# 3.2.1 Comparing the treated and untreated groups

library(ggplot2)

# compare the age distribution
ggplot(data, aes(x = age, color = factor(treat))) +
  stat_ecdf() +
  theme_bw() +
  scale_color_discrete(name = "BMI group",
                       labels = c("BMI < 25", "BMI >= 25"))

# compare the comorbidity_burden distribution
ggplot(data, aes(x = comorbidity_burden, color = factor(treat))) +
  stat_ecdf() +
  theme_bw() +
  scale_color_discrete(name = "BMI group",
                       labels = c("BMI < 25", "BMI >= 25"))

# 3.2.2 ETT computation

etts <- data[, .(ett = mean(cate)), by = treat]
etts
```

We conclude that the age distributions of patients with BMI < 25 and BMI ≥ 25 are fairly similar overall.

We conclude that comorbidity burden appears very similar between the two BMI groups, with no obvious meaningful difference.  

For patients with BMI ≥ 25, the estimated treatment effect is approximately -0.0118, corresponding to about a 1.18 percentage point decrease in mortality probability. For patients with BMI < 25, the estimated treatment effect is approximately -0.0098, corresponding to about a 0.98 percentage point decrease in mortality probability.

```{r 3.3 z-TE(s)}
comorbidity_burden_te <- data[, list(
  zte = mean(cate),
  dev = sd(cate) / sqrt(.N),
  cnt = .N
), by = comorbidity_burden]

ggplot(comorbidity_burden_te[cnt > 10], aes(x = comorbidity_burden, y = zte)) + 
  geom_point() +
  geom_line() +
  geom_ribbon(aes(
    ymin = zte - 1.96 * dev,
    ymax = zte + 1.96 * dev
  ), alpha = 0.3) +
  theme_bw() +
  labs(
    x = "Comorbidity Burden",
    y = "Estimated Effect of BMI on Mortality"
  )

sex_te <- data[, .(
  zte = mean(cate),
  dev = sd(cate) / sqrt(.N),
  cnt = .N
), by = sex]

ggplot(sex_te, aes(x = factor(sex), y = zte)) +
  geom_point() +
  geom_errorbar(aes(
    ymin = zte - 1.96 * dev,
    ymax = zte + 1.96 * dev
  ), width = 0.1) +
  theme_bw() +
  labs(
    x = "Sex (0 = Female, 1 = Male)",
    y = "Estimated Effect of BMI on Mortality"
  )

adm_te <- data[, .(
  zte = mean(cate),
  dev = sd(cate) / sqrt(.N),
  cnt = .N
), by = admission_type]

ggplot(adm_te[cnt > 10], aes(x = reorder(admission_type, zte), y = zte)) +
  geom_point() +
  geom_errorbar(aes(
    ymin = zte - 1.96 * dev,
    ymax = zte + 1.96 * dev
  ), width = 0.1) +
  coord_flip() +
  theme_bw() +
  labs(
    x = "Admission Type",
    y = "Estimated Effect of BMI on Mortality"
  )

decile_heterogeneity <- function(var, data) {
  
  qs <- quantile(data[[var]], seq(0, 1, 0.1), na.rm = TRUE)
  qs <- unique(qs)
  
  data[, dec := .bincode(get(var), qs, include.lowest = TRUE)]
  
  z_te <- data[!is.na(dec), .(
    zte = mean(cate),
    dev = sd(cate) / sqrt(.N),
    cnt = .N
  ), by = dec]
  
  ggplot(z_te[cnt > 5], aes(x = dec, y = zte)) +
    geom_point() +
    geom_line() +
    geom_ribbon(aes(
      ymin = zte - 1.96 * dev,
      ymax = zte + 1.96 * dev
    ), alpha = 0.3) +
    theme_bw() +
    labs(
      x = paste(var, "decile"),
      y = "Estimated effect of BMI on mortality"
    )
}

decile_heterogeneity("age", data)

data[, treat30 := as.integer(bmi_all >= 30)]

crf30 <- causal_forest(
  X = X,
  Y = as.numeric(data$death),
  W = data$treat30
)

mean(predict(crf30)$predictions)
```
















